A Motion Capture and Imitation Learning-based Approach to Robot Control

### 2022-06-28 tex-arxiv/article.tex
fix intro, adjust image spacing again

Manipulator arms and other types of robots have become ubiquitous in modern industry and their use has been proliferating for decades, yet even now the primary method for programming these devices remains procedural code, hand-crafted by specially trained technicians and engineers. This significantly increases the cost and complexity of commissioning process nodes that utilize industrial robots. To this end, various alternative approaches grounded in machine learning have been proposed. Perhaps the most commonly studied are methods that fall under the umbrella of reinforcement learning, which optimize an agent acting on its environment through the use of an explicitly specified reward function \citep{sutton2018reinforcement}. While offering strong theoretical guarantees of convergence, they require interaction with the environment for learning to occur. Furthermore, in practical settings the reward function for many tasks is very sparse, leading to large search spaces and inefficient exploration \citep{hester2018deep}. Finally, the reward function for a task may be unknown or difficult to specify analytically \citep{abbeel2004apprenticeship}.

Imitation learning is a broad field of study in its own right that promises to overcome some or all of the aforementioned challenges, provided that it is possible to obtain a corpus of demonstrations showing how the task is to be accomplished. When dealing with industrial robotics, this is often the case as the tasks we wish to execute are ones that human operators already routinely perform. Therefore, we have developed an approach for recording actions taken by humans in the physical environment, programmatically producing a training data set structured in the form of explicit demonstrations augmented with additional, extrapolated state inputs and control signals, and training artificial neural network models to reproduce the trajectories therein as motion plans.

Given the inherent trade-off between task specificity and required model complexity when it comes to observation data modality and model output, motion capture has been selected as a convenient middle ground. It does not require a simulated environment as approaches that involve virtual reality do \citep{zhang2018deep,dyrstad2018teaching}, obviates any issues that may come with having to learn motions in robot configuration space by operating in Cartesian coordinates, and does not necessitate the use of complex image processing layers as found in models with visual input modalities \citep{liu2018imitation, zhang2018deep}. The main drawbacks of this approach are that it requires motion capture equipment, which is more expensive than conventional video cameras, and Cartesian motion plans need an additional conversion step into joint space trajectories before it is possible to execute them on a real robot. Given the intended application of this system -- programming robots to perform tasks in an industrial environment -- the advantages of compact models, rapid training times and results that are possible to evaluate ahead of time were deemed to outweigh the drawbacks.

To help guide the development process and serve as a means of evaluation, a sample use case was selected -- object throwing with the goal of extending the robot's reach and improving cycle times. In particular, the task of robotically sorting plastic bottles at a recycling plant was to be augmented with a throwing capability. While on the surface the task is almost entirely defined by elementary ballistics that can be programmed explicitly, it serves as a convenient benchmark for robot programming by way of demonstrations. The task was deemed sufficiently non-trivial when considered from the perspective of a Markov decision process or time-series model only given limited information about system state at any given time step, while also being suitable for intuitive evaluation by human observers due to its intuitively straightforward nature. Furthermore, the relationship between release position, velocity and target coordinates provides an obvious way to quantitatively evaluate model performance against training and validation data -- extrapolated throw accuracy.

@@

\begin{equation}
	t, k, n \in \mathbb{N}, \pi_{\theta}:\lbrace {(s_n)_{t-k}^t} \rbrace \rightarrow A
\end{equation}
which, by allowing for input sequences of variable length, may become a function defined on the entire known state history:

@@

\begin{figure}
	\centering
	\fbox{\includegraphics[width=11.6cm]{../img/target_regression.jpg}}
	\caption{Estimation of the throw target coordinates.}
	\label{fig:fig3}
\end{figure}

@@

\newcommand{\reswidth}{15.95cm}

\begin{figure}[!b]
	\centering
	\fbox{\includegraphics[width=\reswidth]{../img/modeltype_comparison.png}}
	\caption{Results -- comparison between model architectures (\emph{Naive} -- refers to the feedforward architecture, also ``naive behavioral cloning''), the best performance attained in each class. Top row metrics -- cosine similarity, Euclidean distance (global); mean position error (stepwise). Bottom row -- mean rotation error (stepwise), mean throw error.}
	\label{fig:res_model_compare}
\end{figure}

\begin{figure}[!hbt]
	\centering
	\fbox{\includegraphics[width=\reswidth]{../img/each_model_best_perf.png}}
	\caption{Results -- best performance for each hyperparameter value. The top row shows feedforward model results, the bottom -- RNN. In both cases, the smaller, newer dataset with less variance in orientation and target coordinates results in better throw evaluations.  Introducing the time signal significantly improves feedforward model performance, whereas for the RNN improved results could be attained at higher learning rates. In the case of the feedforward model, training for too long and making the model too large initially appears to be detrimental, while for the RNN no clear trend was observed in this respect.}
	\label{fig:rnn_naive_best}
\end{figure}

@@

\begin{figure}[!hbt]
	\centering
	\fbox{\includegraphics[width=15.5cm]{../img/ind_rnn_naive.png}}
	\caption{Results -- cosine and throw accuracy metrics for feedforward and recurrent models only varying in the perceptron count parameter (others held constant) -- best 2 on train and validation sets each.}
	\label{fig:ind_rnn_naive}
\end{figure}

@@

\section{Discussion}

>>>>

Manipulator arms and other types of robots have become ubiquitous in modern industry and their use has been proliferating for decades, yet even now the primary method for programming these devices remains procedural code, hand-crafted by specially trained technicians and engineers. This significantly increases the cost and complexity of commissioning process nodes that utilize industrial robots \citep{arents2022smart}. To this end, various alternative approaches grounded in machine learning have been proposed. Many of the methods explored fall under the umbrella of reinforcement learning, which optimize an agent acting on its environment through the use of an explicitly specified reward function \citep{sutton2018reinforcement}. While offering strong theoretical guarantees of convergence, they require interaction with the environment for learning to occur. Furthermore, in practical settings the reward function for many tasks is very sparse, leading to large search spaces and inefficient exploration \citep{hester2018deep}. Finally, the reward function for a task may be unknown or difficult to specify analytically \citep{abbeel2004apprenticeship}.

Imitation learning is a broad field of study in its own right that promises to overcome some or all of the aforementioned challenges, provided that it is possible to obtain a corpus of demonstrations showing how the task is to be accomplished \citep{attia2018global}. When dealing with industrial robotics, this is often the case as the tasks we wish to execute are ones that human operators already routinely perform. Therefore, we have developed an imitation learning-based approach of our own with practical applications in mid. We record actions taken by humans in the physical environment and programmatically produce a training data set structured in the form of explicit demonstrations. These are augmented with additional, indirectly observed state variables and control signals. We then train artificial neural network models to reproduce the trajectories therein as motion plans in Cartesian space.

Given the inherent trade-off between task specificity and required model complexity when it comes to observation data modality and model output, motion capture has been selected as a convenient middle ground. It does not require a simulated environment as approaches that involve virtual reality do \citep{zhang2018deep,dyrstad2018teaching}, obviates any issues that may come with having to learn motions in robot configuration space by operating in Cartesian coordinates, and does not necessitate the use of complex image processing layers as found in models with visual input modalities \citep{liu2018imitation, zhang2018deep}. The main limitations of this approach are that it requires motion capture equipment, which is more expensive than conventional video cameras, and Cartesian motion plans need an additional conversion step into joint space trajectories before it is possible to execute them on a real robot. Given the intended application of this system -- programming robots to perform tasks in an industrial environment -- the advantages of compact models, rapid training times and results that are possible to evaluate ahead of time were deemed to outweigh the drawbacks.

To help guide the development process and serve as a means of evaluation, an example use case was selected -- object throwing with the goal of extending the robot's reach and improving cycle times. In particular, the task of robotically sorting plastic bottles at a recycling plant was to be augmented with a throwing capability. While on the surface the task is almost entirely defined by elementary ballistics that can be programmed explicitly, it serves as a convenient benchmark for robot programming by way of demonstrations. The task was deemed sufficiently non-trivial when considered from the perspective of a Markov decision process or time-series model only given limited information about system state at any given time step, while also being suitable for intuitive evaluation by human observers due to its intuitively straightforward nature. Furthermore, the relationship between release position, velocity and target coordinates provides an obvious way to quantitatively evaluate model performance against training and validation data -- extrapolated throw accuracy.

In this paper, we start by introducing some key concepts in imitation learning -- in particular, how they tie in with the autoregressive use of sequence-to-sequence models \ref{sec:prelim}. Then we explore the broad directions research in this field has taken, as well as prior work relating specifically to the two key practical aspects of our work -- the use of motion capture and throwing tasks \ref{sec:related}. This is followed by a detailed explanation of the reasoning behind our approach \ref{sec:approach}, as well as a description of the practical implementation \ref{sec:materials}. Metrics for evaluating performance on the motivating use case are also introduced in subsection \ref{sec:eval}. In the results \ref{sec:results} and discussion \ref{sec:discussion} sections, we compare the performance of the various model types.

@@

\begin{equation}
	t, k, n \in \mathbb{N}; \pi_{\theta}:\lbrace {(s_n)_{t-k}^t} \rbrace \rightarrow A
\end{equation}
which, by allowing for input sequences of variable length, may become a function defined on the entire known state history:

@@

\begin{figure}
	\centering
	\fbox{\includegraphics[width=13.5cm]{../img/target_regression.jpg}}
	\caption{Estimation of the throw target coordinates.}
	\label{fig:fig3}
\end{figure}

@@

\newcommand{\reswidth}{15.5cm}

\begin{figure}[]
	\centering
	\fbox{\includegraphics[width=\reswidth]{../img/modeltype_comparison.png}}
	\caption{Results -- comparison between model architectures (\emph{Naive} -- refers to the feedforward architecture, also ``naive behavioral cloning''), the best performance attained in each class. Top row metrics -- cosine similarity, Euclidean distance (global); mean position error (stepwise). Bottom row -- mean rotation error (stepwise), mean throw error.}
	\label{fig:res_model_compare}
\end{figure}

\begin{figure}[]
	\centering
	\fbox{\includegraphics[width=\reswidth]{../img/each_model_best_perf.png}}
	\caption{Results -- best performance for each hyperparameter value. The top row shows feedforward model results, the bottom -- RNN. In both cases, the smaller, newer dataset with less variance in orientation and target coordinates results in better throw evaluations.  Introducing the time signal significantly improves feedforward model performance, whereas for the RNN improved results could be attained at higher learning rates. In the case of the feedforward model, training for too long and making the model too large initially appears to be detrimental, while for the RNN no clear trend was observed in this respect.}
	\label{fig:rnn_naive_best}
\end{figure}

@@

\begin{figure}[!hbt]
	\centering
	\fbox{\includegraphics[width=16cm]{../img/ind_rnn_naive.png}}
	\caption{Results -- cosine and throw accuracy metrics for feedforward and recurrent models only varying in the perceptron count parameter (others held constant) -- best 2 on train and validation sets each.}
	\label{fig:ind_rnn_naive}
\end{figure}

@@

\section{Discussion}
\label{sec:discussion}
